---
title: "0008.1-plot-sfs-and-pca"
author: "Mac Campbell"
date: "8/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(gdata)
library(RcppCNPy)
library(viridis)
library(ggrepel)
```

## Plot SFS 
```{r}
sfs<-(as.numeric(read.table("0008/with-out-paralogs.sfs"))) 
barplot(sfs[-c(1,length(sfs))]) #plot variable sites 
```
```{r}
sfs2<-(as.numeric(read.table("0008/with-paralogs.sfs"))) 
barplot(sfs2[-c(1,length(sfs2))]) #plot variable sites 
```


## PCA

Organizing meta.

```{r}
meta<-read_csv("meta/sample-meta.csv")

bams<-read_tsv("bamlists/spineflower.bamlist", col_names=c("Path"))
bams$`NFGEL#`<-gsub("/home/djprince/active_projects/spineflower/1004_alignments/","",bams$Path)
bams$`NFGEL#`<-gsub("_sorted_proper_rmdup.bam","",bams$`NFGEL#`)
meta<-left_join(bams, meta) %>% rename(Label=`NFGEL#`)

counts<-read_csv("meta/spineflower.counts", col_names=c("Counts","SubCounts"))
stats<-read_csv("meta/spineflower.stats", col_names=c("Stat1","Stat2","Mapped"))

meta<-bind_cols(meta,counts,stats)  %>% mutate(Index=1:n())

write_csv(meta, "meta/samples-187.csv")
```

```{r}
ggplot(meta) +
  geom_histogram(aes(x=Counts))

```

Read 187 samples and 29514 sites       
Number of sites after MAF filtering (0.05): 13117        
Estimating admixture with K=11, alpha=0, batch=10, seed=0

```{r}
cov<-read_delim("0008/with-out-paralogs-pca.cov", col_names=FALSE, delim=" ") %>% as.matrix()
```

Masking related individuals with pair-wise kinship estimates >= 0.0625       
Keeping 180 individuals after filtering (removing 7)        
Boolean vector of unrelated individuals saved as 0008/with-out-paralogs-pca-relate.unrelated.npy (Binary)       
Parsing Beagle file       
Read 180 samples and 29514 sites       
Number of sites after MAF filtering (0.05): 13096
Estimating admixture with K=10, alpha=0, batch=10, seed=0

```{r}
covk<-read_delim("0008/with-out-paralogs-pca-relate.cov", col_names=FALSE, delim=" ") %>% as.matrix()
```


```{r}
#' @param samples character vector with the individuals IDs in the order in which
#' they were passed in the bamlist to angsd.
#' @param cov covariance matrix
covar2pcs <- function(samples, cov) {
  
  
  eig <- eigen(cov, symm = TRUE)
  PC <- as.data.frame(eig$vectors) %>%
    as_tibble() %>%
    setNames(sprintf("PC-%02d", 1:ncol(.)))
  
  samtib <- tibble(sample = samples)
  
  list(
    PCs = bind_cols(samtib, PC),
    eigevalues = eig$values
  )
}
```

### Without excluding related inds

```{r}
pca <- covar2pcs(meta$Label, cov)

pca_long <- pca$PCs %>%
  tidyr::gather(., key = "PC", "val", -sample)

# then expand a grid of the possible comparisons (ordered)
expg <- expand.grid(sample = pca$PCs$sample,
                    PCx = sprintf("PC-%02d", 1:6),
                    PCy = sprintf("PC-%02d", 1:6),
                    stringsAsFactors = FALSE) %>%
  tibble::as_tibble()

# then left join the pca results onto that
pca_pairs <- dplyr::left_join(expg, pca_long, by = c("sample", "PCx" = "PC")) %>%
  dplyr::rename(val_x = val) %>%
  dplyr::left_join(pca_long, by = c("sample", "PCy" = "PC")) %>%
  dplyr::rename(val_y = val)

npc <- 3
pc <- pca_pairs %>%
  filter( (PCx %in% sprintf("PC-%02d", 1:npc)) & 
            (PCy %in% sprintf("PC-%02d", 1:npc)) ) %>%
  left_join(., meta, by = c("sample" = "Label")) %>%
  mutate(group = Location) 

eig <- eigen(cov, symm = TRUE)
var <-eig$values/sum(eig$values)
cumvar <-cumsum(eig$values)/sum(eig$values)

head(var)
head(cumvar)
```

```{r}
ggplot(pc) +
  geom_point(aes(x=val_x, y=val_y, color=Location), alpha=0.75) +
  facet_grid(PCx ~ PCy)
```

Examining relatedness.

```{r}
kin<-npyLoad("0008/with-out-paralogs-pca.kinship.npy")

#Convert upper half of matrix to data frame compatible 
kinrind<-which(upper.tri(kin, diag = TRUE), arr.ind = TRUE)
kin<-cbind(kinrind, kin[kinrind])

kindf<-as_tibble(kin) %>%
  rename(Ind1=row, Ind2=col, Kinship=V3) %>%
  filter(Ind1 != Ind2) %>%
  mutate(Mean=mean(Kinship))

#Can I replace Ind1 and Ind2 with individual labels???
df<-kindf %>% left_join(meta, by=c("Ind1"="Index")) %>%
  left_join(meta, by=c("Ind2"="Index")) 

#Ahh, and what are our "related" individuals
df %>% filter(Kinship>0.0625) %>% group_by(Location.x, Location.y) %>% summarize(Count=n())
```

```{r}
popdf<-df %>% filter(Location.x==Location.y) %>% mutate(Site=Location.x)

cols<-viridis(6)

ggplot(popdf)+geom_histogram(aes(x=Kinship), bins=50, fill="grey", color="grey") +
  geom_vline(xintercept =.25, col=cols[1]) + #Parent offspring/full sibs
  geom_vline(xintercept =.125, col=cols[2]) + #2nd degree
  geom_vline(xintercept =.0625, col=cols[3]) + # 3rd
  geom_vline(xintercept = .0313, col=cols[4]) +
  geom_vline(xintercept = .0156, col=cols[5]) +
  geom_vline(xintercept = 0.0, col=cols[6]) +
  scale_x_continuous(breaks=c(0,0.0156, 0.0313, 0.0625, 0.125, 0.25),
                     labels=c("0 / Unrelated","5th Deg.","4th Deg.","3rd Deg.", "2nd Deg.", "1st Deg.")) +
  theme_bw() +
  theme(axis.text.x = element_text(hjust=1, angle=45)) +
  theme(panel.grid=element_blank()) +
  ylab("Count") +
  xlab("Kinship Coefficient")+
  facet_wrap(Site ~ ., ncol=4, scales="free_y")
```

Popmeans
```{r}
popmeans<- df %>% filter(Location.x==Location.y)  %>% group_by(Location.x, Location.y) %>%
  summarize(Mean=mean(Kinship)) %>%
  arrange(-Mean) %>% ungroup() %>% mutate(Position=1:n())

ggplot(popmeans) +
  geom_point(aes(x=Position, y=Mean)) +
  #geom_label_repel(aes(x=Position, y=Mean, label=Location.x)) +
  scale_x_continuous(breaks=popmeans$Position, labels = popmeans$Location.x) +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45)) 
```

### Excluding related inds

```{r}
vec <- npyLoad("0008/with-out-paralogs-pca-relate.unrelated.npy") %>% as_tibble() %>%
  rename(Kept=value)
metak <- bind_cols(meta, vec) %>% filter(Kept !=0)
pcak <- covar2pcs(metak$Label, covk)

pca_longk <- pcak$PCs %>%
  tidyr::gather(., key = "PC", "val", -sample)

# then expand a grid of the possible comparisons (ordered)
expgk <- expand.grid(sample = pcak$PCs$sample,
                    PCx = sprintf("PC-%02d", 1:6),
                    PCy = sprintf("PC-%02d", 1:6),
                    stringsAsFactors = FALSE) %>%
  tibble::as_tibble()

# then left join the pca results onto that
pca_pairsk <- dplyr::left_join(expgk, pca_longk, by = c("sample", "PCx" = "PC")) %>%
  dplyr::rename(val_x = val) %>%
  dplyr::left_join(pca_longk, by = c("sample", "PCy" = "PC")) %>%
  dplyr::rename(val_y = val)

npc <- 3
pck <- pca_pairsk %>%
  filter( (PCx %in% sprintf("PC-%02d", 1:npc)) & 
            (PCy %in% sprintf("PC-%02d", 1:npc)) ) %>%
  left_join(., metak, by = c("sample" = "Label")) %>%
  mutate(group = Location) 

eigk <- eigen(covk, symm = TRUE)
vark <-eigk$values/sum(eigk$values)
cumvark <-cumsum(eigk$values)/sum(eigk$values)

head(vark)
head(cumvark)
```

```{r}
ggplot(pck) +
  geom_point(aes(x=val_x, y=val_y, color=Location), alpha=0.75) +
  facet_grid(PCx ~ PCy)
```

Newhall Ranch is a different species.

```{r}
sub<-pck[!startsWith(pck$Location,"Newhall Ranch"),]
```

```{r}
ggplot(sub) +
  geom_point(aes(x=val_x, y=val_y, color=Location), alpha=0.75) +
  facet_grid(PCx ~ PCy)
```