---
title: "0011-pwfst"
author: "Mac Campbell"
date: "September 1, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message = FALSE)
```

```{r}
library(tidyverse)
library(viridis)
library(ggpubr)
library(cowplot)
library(grid)
library(gridExtra)
```

## Generate pairwise Fst
Need SFS, so see 0011_calc_sfs
See 0011_calc_pwfst.sh    

```{r}
meta<-read_csv("meta/182.csv") %>% separate(Location, into=c("Area","Site"), sep=",", remove=FALSE) 
meta$Site<-trimws(meta$Site)
meta$Site<-gsub(" ","_",meta$Site)
meta<-meta %>%
  mutate(ShortName=ifelse(Area=="Upper Las Virgenes Canyon OSP", "UVLC", Area)) %>%
  mutate(Pop=paste0(ShortName,"-",Site))
```

Making individual pop bamlist

```{r}
poplist<-unique(meta$Pop)
```

```{r}
WriteBamlist <- function (popname) {
 popdf<-filter(meta, Pop==popname) %>% select(Path)
 write_tsv(popdf, paste0("bamlists/",popname,".bamlist"), col_names = FALSE)
}
```

```{r}
lapply(poplist, WriteBamlist)
```


Group all, and then make big bamlists

```{r}
write_tsv(meta %>% filter(Area=="Upper Las Virgenes Canyon OSP") %>% select(Path), "bamlists/UVLC.bamlist", col_names = FALSE)
write_tsv(meta %>% filter(Area=="Newhall") %>% select(Path), "bamlists/Newhall.bamlist", col_names = FALSE)
```

Making a poplist.
```{r}
pops<-c(poplist, "UVLC","Newhall") %>% as_tibble()
write_tsv(pops, file = "meta/poplist", col_names = FALSE)
```

Running 0011_calc_sfs.sh

Running 0011_calc_pwfst.sh

Did something like this with relict dace. Thanks, Iso!   

From the docs:
    
../misc/realSFS fst stats here.fst.idx     
-> FST.Unweight:0.069395 Fst.Weight:0.042349

```{r}
dat<-read_tsv("0011/pwfst.all", col_names = c("Pop1","Pop2","FstUnweighted","FstWeighted")) 
fstdat<-drop_na(dat)

```

```{r}
fstdat$Pop1<-factor(fstdat$Pop1, levels=poplist)
fstdat$Pop2<-factor(fstdat$Pop2, levels=poplist)

```


```{r}
size=6
```

```{r}
fst<-ggplot(drop_na(fstdat), aes(Pop2, Pop1)) + 
  geom_tile(aes(fill = FstWeighted), color="white") + 
  theme_minimal() + 
  scale_fill_viridis(discrete=FALSE, option = "cividis", space = "Lab", direction = -1) + 
  theme_bw() + 
  theme(axis.text.x=element_text(angle=45,vjust=1,size=size,hjust=1), panel.border=element_blank(),
        panel.grid.major=element_blank(), axis.title.y = element_blank(),
        axis.title.x = element_blank()) + scale_y_discrete(position = "left") + 
  theme(axis.text.y=element_text(size=size)) +
 # geom_text(aes(label = round(FstWeighted, 2)), size=2, colour = "white") + 
  ggtitle("A") + 
  theme(plot.title = element_text(face="bold", size=16)) +
  labs(fill = "Fst")

fst + geom_text(aes(label = round(FstWeighted, 2)), size=2, colour = "white") 

ggsave("0011/fstplot.jpg")
```


Following 0011_calc_sfs.sh we can use this command for each pop do generate a sfs.

./misc/realSFS saf2theta out.saf.idx -outname out -sfs out.sfs

(base) maccamp@farm:~/spineflower/0011$ realSFS saf2theta Newhall-Entrada.saf.idx -outname Newhall-Entrada -sfs Newhall-Entrada.sfs

Estimating for each scaffold/chromosome ./misc/thetaStat do_stat out.thetas.idx:    

(base) maccamp@farm:~/spineflower/0011$ thetaStat do_stat Newhall-Entrada.thetas.idx (1283 entries.)

```{sh, eval=FALSE}
cat meta/poplist | while read pop; do realSFS saf2theta 0011/$pop.saf.idx -outname $pop -sfs 0011/$pop.sfs; thetaStat do_stat 0011/$pop.thetas.idx; done;

mv *thetas* ./0011

# we want to combine pestPG files
grep "" 0011/*pestPG | sed 's/.thetas.idx.pestPG:/\t/' | sed 's:0011/::'

| perl -pe 's/_-_/\t/g' > 0011/diversity.all
```

Columns are Pop	Chr	tW	tP	Tajima


```{r}
theta<-read_tsv("0011/diversity.all", col_names = c("Pop","Chr","tW","tP", "Tajima")) %>% filter(tW < 5) %>% filter(tP < 5) %>%
  filter(Pop !="Newhall") %>% filter(Pop != "UVLC")
theta %>% select(-Chr) %>% group_by(Pop) %>% summarize_all(mean)
```


```{r}
pi <- ggplot(theta, aes(x=Pop, y=tP)) + 
  stat_boxplot(geom ='errorbar') +
  geom_boxplot(aes(fill=Pop), notch=FALSE, outlier.size=0.5, outlier.stroke = 0, outlier.alpha=0.5) + 
  labs(y = "Theta Pi") + 
  labs(x="") +
  theme_bw() +
#  scale_y_continuous(breaks=seq(0.00,0.05,0.0025)) + 
  ggtitle(label = "B") +
  theme(plot.title = element_text(face="bold", size=16)) +
  theme(panel.grid = element_blank()) +
  theme(legend.position = "") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1), axis.title.x=element_blank()) +
  theme(axis.text.x = element_text(size=size))

pi
```

```{r}
ggarrange(fst, pi, ncol=2, nrow=1, widths=c(1.5,1))

ggsave("0011/fst-pi.jpg", width=11, height=8.5/2)
```

## Four panel figure

```{r}
load("0009/bic.rda")
load("0009/table.rda")
blank <- grid.rect(gp=gpar(col="white"))
```


```{r}
BIC<- BIC + ggtitle(label = "A") +
  theme(plot.title = element_text(face="bold", size=16)) 
table<-table+theme(axis.title.y=element_blank()) +   ggtitle(label = "B") +
  theme(plot.title = element_text(face="bold", size=16))  +
  theme(axis.title.x = element_text(size=12)) +
  theme(axis.text.y = element_text(face="plain")) +
  theme(legend.position = "")
ab<-ggarrange(arrangeGrob(blank,BIC,blank, ncol=3, widths=c(1,3,1)), blank,table, ncol=1, nrows=3, heights=c(2,.5,5))
ab
ggsave("0011/figure.jpg", width=11/2, height=8.5)
```

```{r}
cd<-ggarrange(fst+ggtitle(label="C"), pi+ggtitle(label="D"), ncol=1, nrow=2)
```

```{r}
ggarrange(ab,cd)
ggsave("0011/figurex.jpg", width=11, height=8.5)
```
